Reasoning About a Machine with Local Capabilities

### 2019-01-24 toplas/paper.tex
Implement reviewer 1 comments

\usepackage{cite}
\usepackage{wrapfig}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{thmtools}
\usepackage{stmaryrd}
% \usepackage[numbers]{natbib}
\usepackage{url}
\usepackage{array}
\usepackage{arydshln}
\usepackage{ifthen}
\usepackage{ifpdf}
\usepackage{verbatim}
\usepackage{mathpartir}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{multicol}
\lstset{
  basicstyle=\footnotesize\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  mathescape
}

@@

\definecolor{OliveGreen}{RGB}{60,128,49}
\definecolor{Blue}{RGB}{92,26,143}
\newenvironment{toplas}{}{}
%    {\color{OliveGreen}}{}
\newenvironment{toplassug}
    {\begin{comment}}{\end{comment}}

@@

\begin{figure}[htb]
  \centering
  \begin{align*}
   \readCond{}(\gl)(W) &=  \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{array}{l}
       \exists r \in \var{localityReg}(g,W) \ldotp \\
\quad   \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{array}
    \right\}\\
   \writeCond{}(\iota,\gl)(W) &=  \left\{
    \npair{(\start,\addrend)}
    \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & W(r) \text{ is address-stratified and} \\[-0.2cm]
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
    \end{aligned} \right\}\\
   \execCond{}(\gl)(W) &= 
    \left\{
      \npairP{(\var{P}, \start,\addrend)}
     \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W, a \in [\start',\addrend'] \subseteq [\start,\addrend], \perm \in \var{P} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start',\addrend',\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
   \entryCond{}(\gl)(W) &= 
    \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{array}{l}
      \forall n' < n \ldotp \forall W' \future W \ldotp\\
      \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{array}
    \right\} \\
   \text{where } \gl = \local \Rightarrow \future = \futurewk \text{ and } \gl = \glob \Rightarrow \future = \futurestr \span
   \end{align*}
\caption{Permission-based conditions}
\label{fig:perm-conds}
\end{figure}
\begin{figure}[htb]
  \begin{align*}
  \iota^\pwl_A, \iota^\nwl_A :{} & \Regions \\
  \iota^\pwl_A \defeq{} & (\temp,1,=,=,H^\pwl_A) \\
  \iota^\nwl_A \defeq{} & (\temp,1,=,=,H^\nwl_A) \\
  \iota^{\nwl,p}_A \defeq{} & (\perma,1,=,=,H^\nwl_A)\\
    H^\pwl_A :{} & \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
  H^\pwl_A \; s \; \hat{W} \defeq{} & \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = A \land \\
      &\forall \addr \in A \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
    \end{aligned}
        \right\} \union \{\npair[0]{\ms} \} \\
    H^\nwl_A :{} & \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
  H^\nwl_A \; s \;\hat{W} \defeq{} & \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = A \land \\
      &\forall \addr \in A \ldotp \\
      &\quad \nonlocal{\ms(\addr)} \land\\ 
      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
    \end{aligned}
        \right\} \union \{\npair[0]{\ms} \}
\end{align*}

@@

The \emph{register-file relation} $\stdrr$ defines safe register-files as those
that contain safe words (i.e.\ words in $\stdvr$) in all registers but $\pcreg$.
The \emph{expression relation} $\stder$ defines that a word is safe to use as a
program counter if it can be plugged into a safe register file (i.e.\ a register
file in $\stdrr$) and paired with a memory satisfying the world to become a safe
configuration. Note that integers and non-executable capabilities (e.g.
$\readonly$ and $\entry$ capabilities) are considered safe program counters
because when they are plugged into a register file and paired with a memory, the
execution will immediately fail, which is safe.

@@

For capabilities with read permission, the $\readCond{}$ ensures that it can
only be used to read safe words, i.e.\ words in the value relation. To guarantee
this, we require that the addressed memory is governed by a region $W(r)$ that
imposes safety as a requirement on the values contained. This safety requirement
is formulated in terms of a standard region $\iota^\pwl_{[\start,\addrend]}$. It simply
requires all the words in the range $[\start,\addrend]$ to be safe, i.e.\ in the
value relation. Requiring that $W(r)\nsubsim[n]\iota^\pwl_{[\start,\addrend]}$
means that $W(r)$ must accept only safe values like
$\iota^\pwl_{[\start,\addrend]}$, but can be even more restrictive if desired. The
read condition also takes into account the locality of the capability because,
generally speaking, global capabilities should only depend on permanent regions.
Concretely, we use the function $\var{localityReg}(\gl,W)$, which projects out
all active (non-revoked) regions when the locality $\gl$ is local, but only the
permanent regions when $\gl$ is global. The definition of the standard region
$\iota^\pwl_{[\start,\addrend]}$ can be found
in \figurename~\ref{fig:std-reg}; it makes use of the isomorphism from
Theorem~\ref{thm:world-existence}.

For a capability with write permission, $\writeCond{}$ must be satisfied for the
capability's range of authority. An adversary can use such a capability to write
any word they can get a hold of, and we can safely assume that they can only get
a hold of safe words, so the region governing the relevant memory must allow any
safe word to be written there. In order to make the logical relation as liberal
as possible, we make this a lower bound of what the region may allow. 
% LB: omitted the following for brevity:
% From another point of view, this lower bound is also an upper bound on the amount of
% trust that can be placed in the contents of memory that an adversary has been
% given write access to. 
For write capabilities, we also have to take into account
the two flavours of write permissions: write and write-local. In the case of
write-local capabilities, the region needs to allow (at least) any safe word to
be written, but in the case of write capabilities, the capability cannot be used
to write local capabilities, so the region only needs to allow safe non-local
values. In the write condition, this is handled by parameterizing it with a
region. For the write-local capabilities the write condition is applied with the
standard region $\iota^\pwl_{[\start,\addrend]}$ that we described previously. For
the write capabilities we use a different standard region
$\iota^\nwl_{[\start,\addrend]}$ which requires that the words in
$[\start,\addrend]$ are non-local and safe. As before, we use
$\var{localityReg}$ to pick an appropriate region based on the capability's
locality. Finally, there is a technical requirement that the region must be
\emph{address-stratified}. Intuitively, this means that if a region accepts two
memory segments, then it must also accept every memory segment ``in between'',
that is every memory segment where each address contains a value from one of the
two accepted memory segments. An interesting property of the write condition is
that they prohibit global write-local capabilities which, as discussed in
\sectionname~\ref{sec:stack-and-return-pointer}, is necessary for any safe use of
local capabilities.
The standard region $\iota^\nwl_{[\start,\addrend]}$ is defined in \figurename~\ref{fig:std-reg}.

@@

\begin{figure}[htb]
  \centering
  \begin{definition}[Malloc Specification]
    \label{spec:malloc}
    $c_\malloc$ satisfies the specification for malloc iff the following
    conditions hold:
    \begin{enumerate}
    \item \label{item:malloc-spec:eg} $c_\malloc = ((\entry,\glob),\_,\_,\_)$
    \item \label{item:malloc-spec:reg} There exists a $\iota_{\malloc,0}$ such that
      \begin{enumerate}
      \item \label{item:malloc-spec:perm} $\iota_{\malloc,0}.v = \perma$
      \item \label{item:malloc-spec:independent} For all $\iota' \futurestr \iota_{\malloc,0}, W,i$ with
        $W(i)=\iota'$, we have that
        \begin{equation*}
          \iota'.H (\iota'.s) (\xi^{-1}(W)) = \iota'.H (\iota'.s) (\xi^{-1}([i \mapsto W(i)]))
        \end{equation*}
      \item \label{item:malloc-spec:correct-exec} For all $\Phi \in \ExecConfs$, $\ms_{\var{footprint}}, \hsframe \in
        \HeapSegments$, $i, n, \size \in \nats$, $w_{\var{ret}} \in \Words$,
        $\iota_\malloc \futurestr \iota_{\malloc,0}$, we have that
        \[
          \begin{array}{l}
            \text{If }
            \memheap = \ms_{\var{footprint}} \uplus \hsframe \land \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land \\
            \quad\memreg(r_1) = \size \land \size \geq 0 \land  \memreg(r_0) = w_{\var{ret}} \land \\
            \quad \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\ %and this needs to change with ABI.
            \text{Then, } \\
            \quad \exists \Phi' \in \ExecConfs \ldotp \exists \ms_{\var{footprint}}', \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
            \qquad \exists j \in \nats \ldotp j > 0 \land \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
            \qquad \quad \Phi \step[j] \Phi' \land \\
            \qquad \quad \memheap[\Phi']=\ms_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
            \qquad \quad \iota_{\malloc}' \futurewk \iota_\malloc \land \\
            \qquad \quad \heapSat[\ms_{\var{footprint}}']{n-j}{[i \mapsto \iota_\malloc']} \land \\
            \qquad \quad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
            \qquad \quad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
            \qquad \quad \size - 1 = e'-b'
          \end{array}
        \]
      \item \label{item:malloc-spec:fail} For all $\Phi \in \ExecConfs$,
        \[
          \begin{array}{l}
            \text{If } (\memreg(r_1) \notin \ints \vee \memreg(r_1) < 0) \land
            \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\
            \text{Then } \exists j \in \nats \ldotp \Phi \step[j] \failed
          \end{array}
        \]
      \end{enumerate}
    \end{enumerate}
  \end{definition}

@@

The macro \texttt{scall
    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} captures
those parts of the calling convention related to actually transferring control
to some adversarial code. Specifically, it pushes the contents of the private
registers, $\overline{r_{\var{priv}}}$, to the stack. Then it pushes the
``restoration'' code to the stack. The restoration code will be executed as the
first thing upon return, and restores the stack pointer and the old program
counter. After the restoration code is pushed to the stack, a copy of the pc is
pushed to the stack after it has been adjusted to point to the first instruction
after the jump. From the stack pointer, a protected return pointer is created by
adjusting it to point to the return code and encapsulating it by restricting its
permission to $\entry$. Next, the stack pointer is restricted to the unused
part, and the unused part of the stack is cleared (as discussed in
Section~\ref{sec:stack-and-return-pointer}). Finally, the non-argument registers
are cleared and we jump to $r$. Upon return after the restoration code has been
executed, the restoration code is popped from the stack and the private words we
pushed to the stack before the call are popped to the private registers. The
implementation of \texttt{scall} is presented in
\figurename~\ref{fig:scall-impl}, and the restoration code is presented in
\figurename~\ref{fig:scall-rest-code}. The implementation of \texttt{scall} uses
some of the macros we present next.

\begin{figure}[tb]
    \centering
\begin{lstlisting}
 $i_1$ = move r_t1 pc
 $i_2$ = lea r_t1 $\var{ofc}$
 $i_3$ = load r_stk r_t1
 $i_4$ = pop pc
\end{lstlisting}
    \caption{The restoration code used in \texttt{scall}. The variable
      $\var{ofc}$ is 5 which is the offset to the address where the old stack
      pointer is stored on the stack.}
    \label{fig:scall-rest-code}
\end{figure}

\begin{figure}[htb]
  \begin{toplas}
    \centering
\begin{lstlisting}
// push private registers to the stack
  push $r_{\var{priv},1}$
  ...
  push $r_{\var{priv},n}$
// push restoration code to the stack
  push $\encode(i_1)$
  push $\encode(i_2)$
  push $\encode(i_3)$
  push $\encode(i_4)$
// push old pc to the stack
  move r_t1 pc
  lea r_t1 $\var{ofc}$ 
  push r_t1
// push stack pointer to the stack
  push r_stk
// set up protected return pointer
  move r_0 r_stk
  lea r_0 $\var{ofc_\var{rec}}$ // -5 is the offset to the first instruction of
the recovery code
  restrict r_0 $\encodePermPair((\local,\entry))$
// restrict stack capability to unused part
  geta r_t1 r_stk
  plus r_t1 r_t1 1
  getb r_t2 r_stk
  subseg r_stk r_t1 r_t2
// clear unused part of the stack
  mclear r_stk
// clear non-argument registers
  rclear $R$
  jmp r
after:
// pop the restore code
  pop r_t1
  pop r_t1
  pop r_t1
  pop r_t1
// pop the private state into approriate registers
  pop $r_{\var{priv},1}$
  ...
  pop $r_{\var{priv},n}$
\end{lstlisting}
    \caption{Implementation of \texttt{scall} $r(r_{\var{args},1},\dots,
      r_{\var{args},m},r_{\var{priv},1},\dots, r_{\var{priv},n})$. The
      restoration code is given in \figurename~\ref{fig:scall-rest-code}. The
      variable $\var{ofc}$ is the offset to the label after, and
      $\var{ofc_\var{rec}} = -5$ which is the offset to the first instruction of
      the activation record. The set $R = \RegName \setminus
      \{\texttt{pc},\texttt{r\_stk},\texttt{r\_0},r,r_{\var{args},1},\dots,r_{\var{args},m}\}$.}
    \label{fig:scall-impl}
  \end{toplas}
\end{figure}

@@

The remaining macros are not directly related to the calling convention, but
they help making the examples in \sectionname~\ref{sec:examples} intelligible.
The macros \texttt{push $r$} and \texttt{pop $r$}
add and remove elements from the stack. The macro \texttt{fetch
    $r$ $\var{name}$} fetches the capability related to $\var{name}$ from the
linking table and stores it in register $r$. The macro
\texttt{malloc $r$ $n$} invokes malloc with size argument $n$.
The \texttt{malloc} macro assumes that a capability for malloc resides in the
linking table and is basically a fetch of the malloc capability followed by a
setup of a return pointer. Finally, the macro \texttt{crtcls
    $\overline{(x_i,r_i)}$ $r$} allocates a closure where $r$ points to the
closure's code and a new environment is allocated (using malloc) and the
contents of $\overline{r_i}$ is stored in the environment. In the code referred
to by $r$, an implicit load from the environment happens when an instruction
refers to $x_i$.

@@

\begin{toplas}
\section{Reasoning about programs on a capability machine}
\label{sec:reasoning}
% \itoplassug{
%   \begin{itemize}
%   \item A section with the lemmas we use for reasoning about programs
%     \begin{itemize}
%     \item lemma 58, scall works. Include description of how it handles the
%       common things in call and leaves the argumentation for callee code and
%       continuation.
%     \item lemma 59, malloc works. If this is included, we should probably also
%       include the malloc specification.
%     \item lemma 60, create closure works.
%     \item Include a description about how the FTLR is used to reason about
%       unknown code.
%     \item Mention that this is a step on the way towards having a program logic,
%       and then in the discussion discuss why a program logic for this language
%       is not what we need.
%     \end{itemize}
%   \end{itemize}
% }
There are many details to get right when programming in assembly. These details
carry over to proofs about assembly programs, so many of the proofs in the next
section about example programs are a bit cumbersome. It is especially annoying,
when the same line of reasoning is applied in multiple places. To mitigate this,
we have defined a number of lemmas that capture common reasoning patterns in
these proofs. We present the most important such lemmas in this section, to give
the reader an idea of how they work.

@@

\subsection{Awkward example}
\textbf{The region for variable $x$} The region $\iota_x$, is the region omitted from the proof sketch for the awkward example.
\figurename~\ref{fig:x-reg} illustrates the transition system of $\iota_x$.
  \begin{figure}[tb]
    \centering
\begin{tikzpicture}[->,shorten >=1pt,auto,node distance=3cm,
  thick,main node/.style={circle,fill=white,draw,font=\sffamily\large},rect node/.style={rectangle,fill=blue!20,draw,font=\sffamily\large}]

>>>>

\usepackage{cite}
\usepackage{wrapfig}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{thmtools}
\usepackage{stmaryrd}
% \usepackage[numbers]{natbib}
\usepackage{xcolor}
\usepackage{url}
\usepackage{array}
\usepackage{arydshln}
\usepackage{ifthen}
\usepackage{ifpdf}
\usepackage{verbatim}
\usepackage{mathpartir}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{multicol}
\lstset{
  basicstyle=\footnotesize\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  mathescape,
  escapeinside=||,
  numbers=left,
  numberblanklines=false,
%  frame=single,
%  framexleftmargin=12pt,
  numbersep=5pt,
  morecomment=[l]{//},
  commentstyle=\color{blue},
  morekeywords={push, pop, move, lea, restrict, subseg, geta, getb, mclear,
    rclear, jmp, load, store, fetch, scall, assert, halt, malloc, call, crtcls,
    reqglob, prepstk, plus, minus, jnz, gete}
}

@@

\definecolor{OliveGreen}{RGB}{60,128,49}
\definecolor{Blue}{RGB}{92,26,143}
\newenvironment{toplas}{}{}
    {\color{OliveGreen}}{}
\newenvironment{toplassug}
    {\begin{comment}}{\end{comment}}

@@

\begin{figure}[htb]
  \centering
  \begin{align*}
   \readCond{}(\gl)(W) &=  \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{array}{l}
       \exists r \in \var{localityReg}(g,W) \ldotp \\
\quad   \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{array}
    \right\}\\
   \writeCond{}(\iota,\gl)(W) &=  \left\{
    \npair{(\start,\addrend)}
    \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & W(r) \text{ is address-stratified and} \\[-0.2cm]
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
    \end{aligned} \right\}\\
   \execCond{}(\gl)(W) &= 
    \left\{
      \npairP{(\var{P}, \start,\addrend)}
     \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W, a \in [\start',\addrend'] \subseteq [\start,\addrend], \perm \in \var{P} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start',\addrend',\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
   \entryCond{}(\gl)(W) &= 
    \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{array}{l}
      \forall n' < n \ldotp \forall W' \future W \ldotp\\
      \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{array}
    \right\} \\
   \text{where } \gl = \local \Rightarrow \future = \futurewk \text{ and } \gl = \glob \Rightarrow \future = \futurestr \span
   \end{align*}
\caption{Permission-based conditions}
\label{fig:perm-conds}
\end{figure}
\begin{figure}[htb]
  \begin{align*}
  \iota^\pwl_A, \iota^\nwl_A :{} & \Regions \\
  \iota^\pwl_A \defeq{} & (\temp,1,=,=,H^\pwl_A) \\
  \iota^\nwl_A \defeq{} & (\temp,1,=,=,H^\nwl_A) \\
  \iota^{\nwl,p}_A \defeq{} & (\perma,1,=,=,H^\nwl_A)\\
    H^\pwl_A :{} & \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
  H^\pwl_A (s)(\hat{W}) \defeq{} & \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = A \text{ and } \\
      &\forall \addr \in A \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
    \end{aligned}
        \right\} \union \{\npair[0]{\ms} \} \\
    H^\nwl_A :{} & \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
  H^\nwl_A (s)(\hat{W}) \defeq{} & \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = A \text{ and } \\
      &\forall \addr \in A \ldotp \\
      &\quad \nonlocal{\ms(\addr)} \text{ and }\\ 
      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
    \end{aligned}
        \right\} \union \{\npair[0]{\ms} \}
\end{align*}

@@

The \emph{register-file relation} $\stdrr$ defines safe register-files as those
that contain safe words (i.e.\ words in the $\stdvr$-relation defined below) in all registers but $\pcreg$.
The \emph{expression relation} $\stder$ defines that a word is safe to use as a
program counter if it can be plugged into a safe register file (i.e.\ a register
file in $\stdrr$) and paired with a memory satisfying the world to become a safe
configuration. Note that integers and non-executable capabilities (e.g.
$\readonly$ and $\entry$ capabilities) are considered safe program counters
because when they are plugged into a register file and paired with a memory, the
execution will immediately fail, which is safe.

@@

For capabilities with read permission, the $\readCond{}$ ensures that it can
only be used to read safe words, i.e.\ words in the value relation. To guarantee
this, we require that the addressed memory is governed by a region $W(r)$ that
imposes safety as a requirement on the values contained. This safety requirement
is formulated in terms of a standard region $\iota^\pwl_{[\start,\addrend]}$. It simply
requires all the words in the range $[\start,\addrend]$ to be safe, i.e.\ in the
value relation. Requiring that $W(r)\nsubsim[n]\iota^\pwl_{[\start,\addrend]}$
means that $W(r)$ must accept only memory segments containing safe values like
$\iota^\pwl_{[\start,\addrend]}$, but can be even more restrictive if needed
(the $\nsubsim[n]$-relation is defined in Appendix~\ref{app:logical-relation}). The
read condition also takes into account the locality of the capability because,
generally speaking, global capabilities should only depend on permanent regions.
Concretely, we use the function $\var{localityReg}(\gl,W)$, which projects out
all active (non-revoked) regions when the locality $\gl$ is local, but only the
permanent regions when $\gl$ is global. The definition of the standard region
$\iota^\pwl_{[\start,\addrend]}$ can be found
in \figurename~\ref{fig:std-reg}; it makes use of the isomorphism $\xi$ from
Theorem~\ref{thm:world-existence}.

For a capability with write permission, $\writeCond{}$ must be satisfied for the
capability's range of authority. An adversary can use such a capability to write
any word they can get a hold of, and we can safely assume that they can only get
a hold of safe words, so the region governing the relevant memory must allow any
safe word to be written there. In order to make the logical relation as liberal
as possible, we make this a lower bound of what the region may allow. 
% LB: omitted the following for brevity:
% From another point of view, this lower bound is also an upper bound on the amount of
% trust that can be placed in the contents of memory that an adversary has been
% given write access to. 
For write capabilities, we also have to take into account
the two flavours of write permissions: write and write-local. In the case of
write-local capabilities, the region needs to allow (at least) any safe word to
be written, but in the case of write capabilities, the capability cannot be used
to write local capabilities, so the region only needs to allow safe non-local
values. In the write condition, this is handled by parameterizing it with a
region. For the write-local capabilities the write condition is applied with the
standard region $\iota^\pwl_{[\start,\addrend]}$ that we described previously. For
the write capabilities we use a different standard region
$\iota^\nwl_{[\start,\addrend]}$ which requires that the words in
$[\start,\addrend]$ are non-local and safe. As before, we use
$\var{localityReg}$ to pick an appropriate region based on the capability's
locality. Finally, there is a technical requirement that the region must be
\emph{address-stratified}. Intuitively, this means that if a region accepts two
memory segments, then it must also accept every memory segment ``in between'',
that is every memory segment where each address contains a value from one of the
two accepted memory segments. An interesting property of the write condition is
that it prohibits global write-local capabilities which, as discussed in
\sectionname~\ref{sec:stack-and-return-pointer}, is necessary for any safe use of
local capabilities.
The standard region $\iota^\nwl_{[\start,\addrend]}$ is defined in \figurename~\ref{fig:std-reg}.

@@

\begin{figure}[htb]
  \centering
  \begin{definition}[Malloc Specification]
    \label{spec:malloc}
    $c_\malloc$ satisfies the specification for malloc iff the following
    conditions hold:
    \begin{enumerate}
    \item \label{item:malloc-spec:eg} $c_\malloc = ((\entry,\glob),\_,\_,\_)$
    \item \label{item:malloc-spec:reg} There exists a $\iota_{\malloc,0}$ such that
      \begin{enumerate}
      \item \label{item:malloc-spec:perm} $\iota_{\malloc,0}.v = \perma$
      \item \label{item:malloc-spec:independent} For all $\iota' \futurestr \iota_{\malloc,0}, W,i$ with
        $W(i)=\iota'$, we have that
        \begin{equation*}
          \iota'.H (\iota'.s) (\xi^{-1}(W)) = \iota'.H (\iota'.s) (\xi^{-1}([i \mapsto W(i)]))
        \end{equation*}
      \item \label{item:malloc-spec:correct-exec} For all $\Phi \in \ExecConfs$, $\ms_{\var{footprint}}, \hsframe \in
        \HeapSegments$, $i, n, \size \in \nats$, $w_{\var{ret}} \in \Words$,
        $\iota_\malloc \futurestr \iota_{\malloc,0}$, we have that
        \[
          \begin{array}{l}
            \text{If }
            \memheap = \ms_{\var{footprint}} \uplus \hsframe \land \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land{} \\
            \quad\memreg(r_1) = \size \land \size \geq 0 \land  \memreg(r_0) = w_{\var{ret}} \land{} \\
            \quad \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\ %and this needs to change with ABI.
            \text{Then, } \\
            \quad \exists \Phi' \in \ExecConfs \ldotp \exists \ms_{\var{footprint}}', \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
            \qquad \exists j \in \nats \ldotp j > 0 \land \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
            \qquad \quad \Phi \step[j] \Phi' \land{} \\
            \qquad \quad \memheap[\Phi']=\ms_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land{}\\
            \qquad \quad \iota_{\malloc}' \futurewk \iota_\malloc \land{} \\
            \qquad \quad \heapSat[\ms_{\var{footprint}}']{n-j}{[i \mapsto \iota_\malloc']} \land{} \\
            \qquad \quad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land{} \\
            \qquad \quad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')} \land{} \\
            \qquad \quad \size - 1 = e'-b'
          \end{array}
        \]
      \item \label{item:malloc-spec:fail} For all $\Phi \in \ExecConfs$,
        \[
          \begin{array}{l}
            \text{If } (\memreg(r_1) \notin \ints \vee \memreg(r_1) < 0) \land
            \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\
            \text{Then } \exists j \in \nats \ldotp \Phi \step[j] \failed
          \end{array}
        \]
      \end{enumerate}
    \end{enumerate}
  \end{definition}

@@

The macro \texttt{scall
    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} captures
those parts of the calling convention related to actually transferring control
to some adversarial code. Specifically, it pushes the contents of the private
registers, $\overline{r_{\var{priv}}}$, to the stack. Then it pushes the
``restoration'' code to the stack. The restoration code will be executed as the
first thing upon return, and restores the stack pointer and the old program
counter. After the restoration code is pushed to the stack, a copy of the pc is
pushed to the stack after it has been adjusted to point to the first instruction
after the jump. From the stack pointer, a protected return pointer is created by
adjusting it to point to the return code and encapsulating it by restricting its
permission to $\entry$. Next, the stack pointer is restricted to the unused
part, and the unused part of the stack is cleared (as discussed in
Section~\ref{sec:stack-and-return-pointer}). Finally, the non-argument registers
are cleared and we jump to $r$. Upon return after the restoration code has been
executed, the restoration code is popped from the stack and the private words we
pushed to the stack before the call are popped to the private registers. The
implementation of \texttt{scall} and the restoration code used by
\texttt{scall} are presented in \figurename~\ref{fig:scall-impl}. The implementation of \texttt{scall} uses
some of the macros we present next.

\begin{figure}[htb]
  \begin{toplas}
    \centering
    \begin{minipage}[t]{.6\linewidth}
\begin{lstlisting}
// Push the private registers to the stack.
  push $r_{\var{priv},1}$
  ...
  push $r_{\var{priv},n}$
// Push the restoration code to the stack.
  push $\encode(i_1)$
  push $\encode(i_2)$
  push $\encode(i_3)$
  push $\encode(i_4)$
// Push the old pc to the stack.
  move r_t1 pc
  lea r_t1 $\var{ofc}$ 
  push r_t1
// Push the stack pointer to the stack.
  push r_stk
// Set up the protected return pointer.
  move r_0 r_stk
  lea r_0 $\var{ofc_\var{rec}}$
  restrict r_0 $\encodePermPair((\local,\entry))$
// Restrict the stack capability to the unused part of the stack.
  geta r_t1 r_stk
  plus r_t1 r_t1 1
  getb r_t2 r_stk
  subseg r_stk r_t1 r_t2
// Clear the unused part of the stack.
  mclear r_stk
// Clear non-argument registers.
  rclear $R$
  jmp r
return:
// Pop the restore code.
  pop r_t1
  pop r_t1
  pop r_t1
  pop r_t1
// Pop the private state into approriate registers.
  pop $r_{\var{priv},1}$
  ...
  pop $r_{\var{priv},n}$
\end{lstlisting}      
    \end{minipage}
    \begin{minipage}[t]{.29\linewidth}
\begin{lstlisting}[title=\footnotesize{The restoration code used in \texttt{scall}.}, captionpos=b,xleftmargin=2em,frame=single,framexleftmargin=1.5em]
 $i_1$ = move r_t1 pc
 $i_2$ = lea r_t1 $\var{ofc}_\var{stk}$
 $i_3$ = load r_stk r_t1
 $i_4$ = pop pc
\end{lstlisting}
    \end{minipage}%
    \caption{Implementation of \texttt{scall} $r(r_{\var{args},1},\dots,
      r_{\var{args},m},r_{\var{priv},1},\dots, r_{\var{priv},n})$. The
      restoration code is presented in the top right corner. The
      variable $\var{ofc}_\var{ret}$ is the offset to the label \texttt{return}, and
      $\var{ofc_\var{rec}} = -5$ which is the offset to the first instruction of
      the activation record. The set $R = \RegName \setminus
      \{\texttt{pc},\texttt{r\_stk},\texttt{r\_0},r,r_{\var{args},1},\dots,r_{\var{args},m}\}$.
    The variable $\var{ofc}_\var{stk} =5$ which is the offset to the address
    where the old stack pointer is stored on the stack.}
    \label{fig:scall-impl}
  \end{toplas}
\end{figure}

@@

The remaining macros are not directly related to the calling convention, but
they help making the examples in \sectionname~\ref{sec:examples} intelligible.
The macros \texttt{push $r$} and \texttt{pop $r$}
add and remove elements from the stack. The macro \texttt{fetch
    $r$ $\var{name}$} fetches the capability related to $\var{name}$ from the
linking table and stores it in register $r$. The macro
\texttt{malloc $r$ $n$} invokes malloc with size argument $n$.
The \texttt{malloc} macro assumes that a capability for malloc resides in the
linking table and is basically a fetch of the malloc capability followed by a
setup of a return pointer. Finally, the macro \texttt{crtcls
    $\overline{(x_i,r_i)}$ $r$} allocates a closure where $r$ points to the
closure's code and a new environment is allocated (using malloc) and the
contents of $\overline{r_i}$ are stored in the environment. In the code referred
to by $r$, an implicit load from the environment happens when an instruction
refers to $x_i$.

@@

\begin{toplas}
\section{Reasoning about programs on a capability machine}
\label{sec:reasoning}
% \itoplassug{
%   \begin{itemize}
%   \item A section with the lemmas we use for reasoning about programs
%     \begin{itemize}
%     \item lemma 58, scall works. Include description of how it handles the
%       common things in call and leaves the argumentation for callee code and
%       continuation.
%     \item lemma 59, malloc works. If this is included, we should probably also
%       include the malloc specification.
%     \item lemma 60, create closure works.
%     \item Include a description about how the FTLR is used to reason about
%       unknown code.
%     \item Mention that this is a step on the way towards having a program logic,
%       and then in the discussion discuss why a program logic for this language
%       is not what we need.
%     \end{itemize}
%   \end{itemize}
% }
There are many details to get right when programming in assembly. These details
carry over to proofs about assembly programs, so many of the proofs in the next
section about example programs are a bit cumbersome. It is especially annoying
when the same line of reasoning is applied in multiple places. To mitigate this,
we have defined a number of lemmas that capture common reasoning patterns in
these proofs. We present the most important such lemmas in this section, to give
the reader an idea of how they work.

@@

\subsection{Awkward example}
\label{app:awkward-example}
\textbf{The region for variable $x$} The region $\iota_x$, is the region omitted from the proof sketch for the awkward example.
\figurename~\ref{fig:x-reg} illustrates the transition system of $\iota_x$.
  \begin{figure}[tb]
    \centering
\begin{tikzpicture}[->,shorten >=1pt,auto,node distance=3cm,
  thick,main node/.style={circle,fill=white,draw,font=\sffamily\large},rect node/.style={rectangle,fill=blue!20,draw,font=\sffamily\large}]
